Thesis: draft

### 2021-03-31 draft.tex
changed some comments
implemented lay-out feedback from Yolan

\documentclass[11 pt , letterpaper , oneside , openright]{book}
\renewcommand{\baselinestretch}{1.15} 
\usepackage[utf8]{inputenc}

@@

@@

{\large \today}\\[4cm] % Date
%\includegraphics[scale=0.7]{ugent_logo.png} % University/department logo - uncomment to place it
% https://www.overleaf.com/project/605b4c033a439c5f0f15c7e5 --> copywright?
\vfill
\end{center}
\end{titlepage}

\pagestyle{plain}

\chapter*{Acknowledgments} 
\addcontentsline{toc}{chapter}{Acknowledgments}

\selectlanguage{english}
\begin{abstract}
\addcontentsline{toc}{chapter}{Abstract}
The past decade, the rise of social media has become undeniable. More and more people are getting engaged on some type of social media platform and hence, it is probably not wrong to say that nowadays most of us have, beside our real life, a virtual life on social media. This rapidly increased engagement on social media undoubtedly has many advantages. However, every upside has a downside and social media is no different. This thesis deals with the impact of social media on the formation and evolution of opinions in society.\\ 
Social media allows for the spread of information at a faster pace and at an unprecedented scale. However, people are still bounded by time- and cognitive constraints. In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to. This filtering and individualizing of information may have some potentially dangerous downfalls in the sense that people might get trapped in "echo chambers" that confirm their own beliefs. It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
People are, however, more than sheep that thoughtlessly follow their shepherd (contrary to what some books might lead us to believe). They are capable of critical thinking and may be hesitant to change their current opinion. This aspect of potential stubbornness should not be ignored.\\
In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions. We will investigate whether network properties such as community structure and clustering coefficient may enhance or hamper the formation of echo chambers. We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness. Beside that, we will also pay some attention to the possible ways of distributing the opinions; we will compare randomly distributed cases to cases where entire communities are given one single opinion.
% definetely not finished!! --> add some type of conclusions or not? talk about running model on real world networks?
\end{abstract}

\selectlanguage{dutch}
\begin{abstract}
\addcontentsline{toc}{chapter}{Samenvatting}
% definetely not finished!! --> this will be put in summary! (think about how you want lay-out: chapter of abstract style?)
\end{abstract}

@@

% add research question and hypotheses!!
% also talk about algorithmic personalisation!!
% ADD MORE REFERENCES (previous research, contextualization etc)

The past decade social media has become ever more important in our daily lives. Nowadays most of us have, beside a real-world life, a virtual life on social media. This increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale. \\
It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings \cite{Perra2019}. The broad range of interactions people undergo on social media with people from all over the world can thus not be underestimated in the process of opinion formation. The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.\\
\newline
It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion. One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
\newline
One important feature of opinion dynamics on social media is the use of algorithmic personalization. Social media allow for simultaneous or asynchronous interactions between people without geographical constraints and thus allow for the spread of information at a faster pace and at an unprecedented scale \cite{Perra2019}. People are, however, still bound by temporal and cognitive constraints. In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}. It is of great importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\  
\newline
Opinion dynamics models have two important layers. On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.\\ Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks. The second layer, the opinion dynamics layer, is introduced in Section \ref{Opinion}. Section \ref{statPhys} introduces concepts such as statistical and social physics. Finally, in Section \ref{goal}, the hypotheses and goals of this thesis are discussed.

@@

One of the goals of this thesis is to investigate/determine the role of the underlying network structure on the formation and dynamics of opinions in the system. Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model, the Watts-Strogatz model, etc. These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}. %references to chapter 2 need to be adapted

@@

% maybe also talk about degree distributions (poisson, power law), power law behavior of many real systems, clustering + maybe talk about networks that will be used in this thesis?

@@

As stated before, the formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg. stubbornness), the information someone is exposed to, etc.  Opinion dynamics models should try to capture these complex processes in a simplified way. Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules. These models are obviously a simplification of real-world opinion dynamics. It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}. \\
\newline
The opinion of the agents in the model can be either discrete or continuous. This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions. This might come across as a huge simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg. republicans or democrats, cat or dog, renting or buying a house,...). \\
\newline
The rules that determine how an agent updates his or her opinion can include many different aspects. They can, for example, be as simple as a majority rule (i.e. the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating. A small overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}. \\
One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others. Furthermore, the agents can have a parameter that determines whether they are active or not. The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
\newline
Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms. The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}. Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}. In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.
% add reference to chapter 2

% not finished, describe different models
% talk about updating rules used in thesis + concepts as resistance etc. 
% biblio: https://link.springer.com/article/10.1007/s13538-020-00772-9

@@

The adjacency matrix contains a lot of information such as the degree of a node. The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of node $i$. The degree can be easily obtained from the adjacency matrix \cite{Mata2020}
\begin{equation}
	k_i = \sum_{j=1}^N A_{ij}
\end{equation}
In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\text{in}}$ and the number of outgoing edges $k_i^{\text{out}}$ of node $i$. These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
\begin{equation}
	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}
\end{equation}
The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$. For weighted graphs the degree is easily generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
\begin{equation}
	s_i = \sum_{j=1}^N w_{ij}
\end{equation}
From the degree of the nodes in the network one can construct the degree distribution. The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$. The average degree can be obtained from the degree distribution \cite{Mata2020}
\begin{equation}
	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k)
\end{equation}
where $N$ is the total number of nodes in the network. The degree distribution also allows us to classify networks. The two most important classes are homogeneous and heterogeneous networks. Homogeneous networks have a bell curved degree distribution, eg. a Poisson distribution. In this case, most of the nodes have a degree close to the average degree $\left<k\right>$. Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$. % explain differences  more, consequences of power-law + need to include figure?
They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}. % correct statement? + talk about fact that av degree can diverge; explain scale free concept a bit more detailed
%maybe explain why we care about degree, centrality measure etc

@@

\subsection{Network distances}

% explain length, diameter, average length,...
Several very useful measurements quantify the notion of length and distance in a network/graph. We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}. The length of a path is then defined as the number of edges in the path \cite{Goddard2010}. A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (note that the geodesic path is not necessarily unique) \cite{F.Costa2007}. The length of the geodesic path between the two nodes $i$ and $j$ is called the geodesic distance or shortest distance $d_{ij}$ \cite{F.Costa2007}. Geodesic distance is often just called distance \cite{Goddard2010}. The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\  
\newline
\newline
\textbf{Average distance}\\
\newline
The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
\begin{equation}\label{avdist}
	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
\end{equation}
This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}. This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes. These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}. Instead, another definition of average distance is used: the global efficiency \cite{F.Costa2007}
\begin{equation}
	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}
\end{equation}
This measurement quantifies the efficiency of the network in sending information between nodes \cite{F.Costa2007}. This assumes that the efficiency of sending information between two nodes is reciprocal to their distance. The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency \cite{F.Costa2007}
\begin{equation}
 	h = \frac{1}{E}
\end{equation}
\newline
\textbf{Diameter and radius}\\
\newline
The diameter $diam(G)$ of a network $G$ is the maximum distance between any pair of nodes \cite{Goddard2010}. The eccentricity of a node is the maximum distance from that node to any other node. The radius $rad(G)$ is the minimum eccentricity among all nodes of $G$. Note that the diameter is the maximum eccentricity among all nodes in $G$ \cite{Goddard2010}.
% finished?

\subsection{Clustering coefficient and transitivity}

@@

	\bar{C} = \frac{1}{N}\sum_{i = 1}^N C_i

@@

It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$. It is then obvious that the local clustering coefficient $C_i$ (Eq. (\ref{clus})) of a node and the transitivity $T_i$ (Eq. (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}. In the case of the average clustering coefficient $\bar{C}$ and the global transitivity $T$ this is, however, not the case. The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node. This may lead to slightly different values since nodes with a higher degree may possibly be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.

@@

Network communities are another feature that (complex) networks may posses. A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}. More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}. In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}. The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}. The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it. Several community definitions are consistent with the density hypothesis. Let us consider three possible definitions: maximum cliques, strong communitoes and weak communities \cite{Albert2016}.\\
\newline
\textbf{Maximum cliques}\\
\newline
A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}. A community based on this definition would then be the largest clique in the network. This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\   
\newline
\textbf{Strong communities}\\
\newline
A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}. Let us denote a community as $C$ and let $i$ be a node inside the community $C$. If we define the internal degree $k_i^{int}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{ext}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
\begin{equation}
	k_i^{int}(C) > k_i^{ext}(C) \qquad \forall i \in C
\end{equation}
\newline
\textbf{Weak communities}\\
\newline
A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}. We thus have the following condition for a weak community $C$ \cite{Albert2016}
\begin{equation}
	\sum_{i \in C} k_i^{int}(C) > \sum_{i \in C}k_i^{ext}(C)
\end{equation}

@@

% talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
% explain small-world (milgram), high clustering, power-law

\section{Network models}
\label{netModel}

@@

	\bar{C} = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}

@@

This can be easily understood since choosing three nodes out of $N$ is given by $\binom{N}{3}$. The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates). If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\bar{C} \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}. \\

@@

The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic/random nature. In this model the fraction of neighbors of user $i$ with opinion A and the fraction of neighbors with opinion B are determined. Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}. Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}. 
% https://arxiv.org/pdf/0710.3256.pdf read this (overview, discusses several models)!!

@@

% talk about epsilon parameter in deffuant model to implement stubbornness + talk about model thesis, binary so no real extreme/moderate distinction

% maybe also mention here the model with the threshold!!

@@

% NOT FINISHED: talk about shortcommings --> models, neither of them describes real networks completely accurate, etc

>>>>

\documentclass[11 pt , letterpaper , twoside , openright]{book}
\renewcommand{\baselinestretch}{1.15} 
\usepackage[utf8]{inputenc}

@@

\usepackage{geometry}
\geometry{
 a4paper,
 left=30mm,
 bottom=25mm,
 right = 20mm
}

\usepackage{emptypage}

@@

{\large Academic year 2020-2021}\\[4cm] % Date
%\includegraphics[scale=0.7]{ugent_logo.png} % University/department logo - uncomment to place it
% https://www.overleaf.com/project/605b4c033a439c5f0f15c7e5 --> copywright?
\vfill
\end{center}
\end{titlepage}

\pagestyle{plain}
\renewcommand{\abstractname}{Acknowledgments}
\begin{abstract}
\addcontentsline{toc}{chapter}{Acknowledgments}

\end{abstract}

\renewcommand{\abstractname}{Abstract}
\selectlanguage{english}
\begin{abstract}
\addcontentsline{toc}{chapter}{Abstract}
The past decade, the rise of social media has become undeniable. More and more people are getting engaged on some type of social media platform and hence, it is probably not wrong to say that nowadays most of us have, beside our real life, a virtual life on social media. This rapidly increased engagement on social media undoubtedly has many advantages. However, every upside has a downside and social media is no different. This thesis deals with the impact of social media on the formation and evolution of opinions in society.\\ 
Social media allows for the spread of information at a faster pace and at an unprecedented scale. However, people are still bounded by time- and cognitive constraints. In order to handle the information overflow, social media companies design filtering algorithms that can fine-tune and individualize the information someone is exposed to. This filtering and individualizing of information may have some potentially dangerous downfalls in the sense that people might get trapped in "echo chambers" that confirm their own beliefs. It is important to know the effects of such algorithmic personalization on the appearance and formation of phenomena such as polarization, spread of extremism, formation of echo chambers, etc.\\
People are, however, more than sheep that thoughtlessly follow their shepherd (contrary to what some books might lead us to believe). They are capable of critical thinking and may be hesitant to change their current opinion. This aspect of potential stubbornness should not be ignored.\\
In this thesis toy models are designed to investigate the interplay between network structure, filtering algorithms and individual resistance to change in the evolution of two competing opinions. We will investigate whether network properties such as community structure and clustering coefficient may enhance or hamper the formation of echo chambers. We will use some basic filtering algorithms to determine their possible effects on the appearance of echo chambers and we will investigate the impact of individual stubbornness. Beside that, we will also pay some attention to the possible ways of distributing the opinions; we will compare randomly distributed cases to cases where entire communities are given one single opinion.
% definetely not finished!! --> add some type of conclusions or not? talk about running model on real world networks?
\end{abstract}

\selectlanguage{dutch}
\begin{abstract}
\addcontentsline{toc}{chapter}{Samenvatting}
% definetely not finished!!
\end{abstract}

@@

% add research question and hypotheses!!
% ADD MORE REFERENCES (previous research, contextualization etc)

The past decade social media has become ever more important in our daily lives. Nowadays most of us have, beside a real-world life, a virtual life on social media. This increased engagement on social media may have a huge impact on the formation of opinions, not only on the individual scale, but also on a broader, societal scale. \\
It is known that individuals not only form their opinions through self-reflection, but also through interactions with other people and with their surroundings \cite{Perra2019}. The broad range of interactions people undergo on social media with people from all over the world can thus not be underestimated in the process of opinion formation. The understanding of the role of social media on the emergence of polarization and extremism in society is of uttermost importance.\\
\newline
It is also known that some people change their opinions more easily than others, that is, once in a while you encounter somebody that is really stubborn and persistent towards its own opinion. One of the aims of this thesis is to design toy models of opinion dynamics with stubborn actors on theoretical and real-world social networks to analyze the interplay between individual resistance to change and the network structure in the evolution of two competing opinions.\\
\newline
One important feature of opinion dynamics on social media is the use of algorithmic personalization. Social media allow for simultaneous or asynchronous interactions between people without geographical constraints and thus allow for the spread of information at a faster pace and at an unprecedented scale \cite{Perra2019}. People are, however, still bound by temporal and cognitive constraints. In order to assure a more pleasant/convenient experience on their platforms, social media companies use algorithms to order/filter the posts on an individual's time-line according to what might be relevant to that individual \cite{Perra2019}. It is of great importance to know the possible effects of these filtering algorithms on the evolution of opinions in the population.\\  
\newline
Opinion dynamics models have two important layers. On one hand, we introduce social networks to describe the underlying structure of social interactions; on the other hand, we need appropriate opinion dynamics/formation models to reproduce the opinion dynamics and opinion formation processes found in real life.\\
\newline
Section \ref{complNet} gives an introduction to the first layer, namely that of complex networks. The second layer, the opinion dynamics layer, is introduced in Section \ref{Opinion}. Section \ref{statPhys} introduces concepts such as statistical and social physics. Finally, in Section \ref{goal}, the hypotheses and goals of this thesis are discussed.

@@

One of the goals of this thesis is to investigate/determine the role of the underlying network structure on the formation and dynamics of opinions in the system. Some examples of theoretical network models used in this thesis are the Erd\H{o}s-R\'{e}nyi network, the stochastic block model, the Watts-Strogatz model, etc. These will be explained thoroughly in Chapter \ref{chap2}, Section \ref{netModel}. 

@@

@@

As stated before, the formation of an individual's opinion is the result of the interplay of many factors such as self-reflection, peer pressure, the individual's personality (eg. stubbornness), the information someone is exposed to, etc.  Opinion dynamics models should try to capture these complex processes in a simplified way. Several models have been developed ranging from simple binary models to more complex, continuous approaches \cite{Sirbu2016}.\\
The basic idea of all opinion dynamics models is that the nodes or agents in a social network have a variable that represents their opinion and that is updated according to some predefined rules. These models are obviously a simplification of real-world opinion dynamics. It is however shown that they do display a lot of aspects of real opinion formation such as agreement, transitions between order (consensus) and disorder (fragmentation), polarization, formation of echo chambers (clusters of people with the same opinion), etc \cite{Sirbu2016}. \\
\newline
The opinion of the agents in the model can be either discrete or continuous. This thesis deals with models where each agent can have one of the two opinions A or B and thus only deals with the case of discrete opinions. This might come across as a huge simplification of the real-world complexity of opinion formation, but also in real life people often have to choose between two competing opinions (eg. republicans or democrats, cat or dog, renting or buying a house,...). \\
\newline
The rules that determine how an agent updates his or her opinion can include many different aspects. They can, for example, be as simple as a majority rule (i.e. the majority model: if the majority of your neighbors have a certain opinion, you adopt that opinion as well) or can include a more probabilistic way of updating. A small overview of possible opinion dynamics models is given in Chapter \ref{chap3}, Section \ref{Bin} and Section \ref{Con}. \\
One can also include a resistance parameter that determines the hesitation of an agent to change to a new opinion and/or a parameter that determines the influence of an agent on others. Furthermore, the agents can have a parameter that determines whether they are active or not. The introduction of stubborn people will be discussed in Chapter \ref{chap3}, Section \ref{stubb}; the activation mechanism of nodes is discussed in Chapter \ref{chap3}, Section \ref{actMech}.\\
\newline
Since this thesis deals with the particular case of opinion dynamics on on-line social platforms, it is important to define/set up different filtering algorithms. The filtering algorithms that real social media companies use, are corporate secrets, but we know that there are three main principles of content curation: popularity, semantic and collaborative filtering \cite{Perra2019}. Popularity filtering refers to the practice of promoting content that is popular across the platform; semantic filtering means that post similar to previous consumed posts are recommended and collaborative filtering suggests posts that are similar to the ones our friends consume \cite{Perra2019}. In Chapter \ref{chap3}, Section \ref{filter} a detailed description of the filtering algorithms used in  this thesis will be given.

% biblio: https://link.springer.com/article/10.1007/s13538-020-00772-9

@@

The adjacency matrix contains a lot of information such as the degree of a node. The degree $k_i$ of a node $i$ is the number of edges attached to node $i$ or, in others words, $k_i$ represents the number of nearest neighbors of node $i$. The degree can be easily obtained from the adjacency matrix \cite{Mata2020}
\begin{equation}
	k_i = \sum_{j=1}^N A_{ij}
\end{equation}
In the case of a directed graph one can differentiate between the number of incoming edges $k_i^{\textrm{in}}$ and the number of outgoing edges $k_i^{\textrm{out}}$ of node $i$. These are called the in-degree and out-degree respectively and are defined as \cite{Mata2020}
\begin{equation}
	k_i^{\text{in}} = \sum_{j=1}^N A_{ji} \quad \text{and} \quad k_i^{\text{out}} = \sum_{j=1}^N A_{ij}
\end{equation}
The total degree of node $i$ is then $k_i = k_i^{\text{in}} + k_i^{\text{out}}$. For weighted graphs the degree is easily generalized to the weighted degree $s_i$, often called strength \cite{Ioannis2007}
\begin{equation}
	s_i = \sum_{j=1}^N w_{ij}
\end{equation}
From the degree of the nodes in the network one can construct the degree distribution. The degree distribution $P(k)$ represents the fraction of nodes with a degree $k$ or, in other words, the degree distribution gives the probability that a randomly selected node has a degree $k$. The average degree can be obtained from the degree distribution \cite{Mata2020}
\begin{equation}
	\left<k\right> = \frac{1}{N} \sum_{i=1}^N k_i = \sum_k k P(k)
\end{equation}
where $N$ is the total number of nodes in the network. The degree distribution also allows us to classify networks. The two most important classes are homogeneous and heterogeneous networks. Homogeneous networks have a bell curved degree distribution, eg. a Poisson distribution. In this case, most of the nodes have a degree close to the average degree $\left<k\right>$. Heterogeneous networks, on the other hand, have a power-law degree distribution, $P(k) \sim k^{-\gamma}$. % explain differences  more, consequences of power-law + need to include figure?
They are also referred to as scale-free networks, since they do not posses a characteristic length scale or, in this case, the average degree isn't a characteristic scale for the network \cite{Mata2020}. % correct statement? + talk about fact that av degree can diverge; explain scale free concept a bit more detailed
%maybe explain why we care about degree, centrality measure etc

@@

\newpage
\subsection{Network distances}

% explain length, diameter, average length,...
Several very useful measurements quantify the notion of length and distance in a network/graph. We start with the definition of a path: a path is a sequence of distinct nodes, such that adjacent nodes in the sequence are adjacent nodes in the network \cite{Goddard2010}. The length of a path is then defined as the number of edges in the path \cite{Goddard2010}. A geodesic path, often called the shortest path, between two nodes is the path between two nodes $i$ and $j$ with minimum length (note that the geodesic path is not necessarily unique) \cite{F.Costa2007}. The length of the geodesic path between the two nodes $i$ and $j$ is called the geodesic distance or shortest distance $d_{ij}$ \cite{F.Costa2007}. Geodesic distance is often just called distance \cite{Goddard2010}. The notion of distance enables to define several network measurements such as average distance, diameter and radius.\\  
\newline
\newline
\textbf{Average distance}\\
\newline
The average distance is the mean value of the geodesic distance $d_{ij}$ \cite{F.Costa2007}
\begin{equation}\label{avdist}
	l = \frac{1}{N(N-1)} \sum_{i \neq j} d_{ij}
\end{equation}
This definition, however, diverges if there are unconnected nodes in the network (unconnected nodes have, by definition, a distance equal to infinity) \cite{F.Costa2007}. This problem could be avoided by only including connected pairs in the sum, but this introduces a distortion for networks with a lot of unconnected pairs of nodes. These networks will then have a small average distance, which is only expected for networks with lots of connections \cite{F.Costa2007}. Instead, another definition of average distance is used: the global efficiency \cite{F.Costa2007}
\begin{equation}
	E = \frac{1}{N(N-1)} \sum_{i \neq j} \frac{1}{d_{ij}}
\end{equation}
This measurement quantifies the efficiency of the network in sending information between nodes \cite{F.Costa2007}. This assumes that the efficiency of sending information between two nodes is reciprocal to their distance. The harmonic mean of the geodesic distances is now defined as the reciprocal of the global efficiency \cite{F.Costa2007}
\begin{equation}
 	h = \frac{1}{E}
\end{equation}
\newline
\textbf{Diameter and radius}\\
\newline
The diameter $diam(G)$ of a network $G$ is the maximum distance between any pair of nodes \cite{Goddard2010}. The eccentricity of a node is the maximum distance from that node to any other node. The radius $rad(G)$ is the minimum eccentricity among all nodes of $G$. Note that the diameter is the maximum eccentricity among all nodes in $G$ \cite{Goddard2010}.
% finished?
\newpage
\subsection{Clustering coefficient and transitivity}

@@

	\left<C\right> = \frac{1}{N}\sum_{i = 1}^N C_i

@@

It is not too hard to see that $N_{\Delta}(i)$ counts the number of edges between the neighbors of $i$ and that $N_3(i)$ is equal to $k_i(k_i-1)/2$, where $k_i$ is the degree of node $i$. It is then obvious that the local clustering coefficient $C_i$ (Eq. (\ref{clus})) of a node and the transitivity $T_i$ (Eq. (\ref{trans})) of a node define the same quantity \cite{F.Costa2007}. In the case of the average clustering coefficient $\left<C\right>$ and the global transitivity $T$ this is, however, not the case. The difference between the two definitions is that the average of Equation (\ref{globalTrans}) gives the same weight to each triangle in the network, whereas Equation (\ref{avClus}) gives the same weight to each node. This may lead to slightly different values since nodes with a higher degree may possibly be involved in a higher number of triangles than nodes with a lower degree \cite{F.Costa2007}.

@@

Network communities are another feature that (complex) networks may posses. A community inside a network is loosely defined as a set of nodes that are more densely connected to nodes inside that set than to the other nodes in the network \cite{Saha2015}. More strictly, a community is defined based on two hypotheses: the connectedness hypothesis and the density hypothesis \cite{Albert2016}. In short, the connectedness hypothesis means that each member of a community should be reached through each other member of the same community \cite{Albert2016}. The density hypothesis implies that nodes inside a community are more likely to be linked to other nodes inside that community than to nodes outside the community \cite{Albert2016}. The density hypothesis narrows what could be considered a community, but it doesn't uniquely define it. Several community definitions are consistent with the density hypothesis. Let us consider three possible definitions: maximum cliques, strong communitoes and weak communities \cite{Albert2016}.\\
\newline
\textbf{Maximum cliques}\\
\newline
A clique is a complete subgraph, where a complete subgraph is defined as a set of nodes of the network where each node in the set is directly connected to all the others nodes in the same set \cite{Albert2016}. A community based on this definition would then be the largest clique in the network. This definition of community might, however, be too restrictive and, beside that, large cliques don't appear very frequent in networks \cite{Albert2016}.\\   
\newline
\textbf{Strong communities}\\
\newline
A strong community is defined such that each node inside the community has more edges to other nodes inside the same community than to nodes outside the community \cite{Albert2016}. Let us denote a community as $C$ and let $i$ be a node inside the community $C$. If we define the internal degree $k_i^{\textrm{int}}(C)$ as the number of edges that connect node $i$ with other nodes in $C$ and the external degree $k_i^{\textrm{ext}}(C)$ as the number of edges that connect node $i$ with nodes outside $C$, then we have the following condition for $C$ to be a strong community \cite{Albert2016}
\begin{equation}
	k_i^{\textrm{int}}(C) > k_i^{\textrm{ext}}(C) \qquad \forall i \in C
\end{equation}
\newline
\textbf{Weak communities}\\
\newline
A weak community is a community where the sum of the internal degree of all the nodes in the community exceeds the sum of the external degree of all the nodes in the community \cite{Albert2016}. We thus have the following condition for a weak community $C$ \cite{Albert2016}
\begin{equation}
	\sum_{i \in C} k_i^{\textrm{int}}(C) > \sum_{i \in C}k_i^{\textrm{ext}}(C)
\end{equation}

@@

% talk about what is a graph, talk about degree distribution, clustereing, average path length, etc (in different subsections), maybe things like assortativity, page rank,...
% explain small-world (milgram), high clustering, power-law
\newpage
\section{Network models}
\label{netModel}

@@

	\left<C\right> = \frac{\binom{N}{3}p^3}{\binom{N}{3}p^2} = p = \frac{\left<k\right>}{N-1}

@@

This can be easily understood since choosing three nodes out of $N$ is given by $\binom{N}{3}$. The chance that they are connected by three edges is given by $p^3$, which gives the number of triangles; the chance that the are connected by two edges is $p^2$, which determines the number of connected triples (recall that this model creates random networks with no self-loops and no edge duplicates). If the average degree $\left<k\right>$ is constant, the average clustering constant goes as $\left<C\right> \sim \mathcal{O}(1/N)$, which becomes very small for large networks \cite{Clauset2011}. \\

@@

The probabilistic majority model is similar to the majority model, but has, as the name suggests, a more probabilistic/random nature. In this model the fraction of neighbors of user $i$ with opinion A and the fraction of neighbors with opinion B are determined. Then, the opinion of user $i$ is updated: he/she takes on opinion A (or B) with a probability equal to the fraction of his/her neighbors with opinion A (or B) \cite{Perra2019}.\\
\newpage
\noindent
For example, if $70\%$ of the neighbors of user $i$ have an opinion A (and thus $30 \%$ of the neighbors have opinion B), user $i$ will update his/her opinion to opinion A with a probability 0.7 and to opinion B with a probability 0.3 \cite{Perra2019}. Thus, user $i$ no longer simply updates his/her opinion according to the majority of opinions carried by his/her neighbors, but the majority opinion still has a higher probability to be taken on by this user.\\
This model is no longer deterministic, but has a level of randomness, which is an important component of opinion dynamics \cite{Perra2019}. 
% https://arxiv.org/pdf/0710.3256.pdf read this (overview, discusses several models)!!

@@

@@
